Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-06-08 main.tex
1 pass section parsing

\item[--] a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
defining a measure $A: \T(\Omega) \to \Semiring$,
%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
%      $A : \Omega^* \to \Semiring$,
\item[--] an input word $s \in \Sigma^*$.
\end{description}
%
As explained in Section~\ref{sec:trees}, 
$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
or a labeled weighted CF grammar whose derivation trees are in $\T(\Omega)$, 
with weight defined by $A$.
%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
%according to Definition~\ref{def:SWTA}, 
%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
%
%This actually is a slight generalization of CFG since 
%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
%hence parse trees %derivation trees
%are trees of $\T(\Omega)$.
%
%Another (more original) generalization is that the set of terminal symbols 
%$\Omega_0$ may be infinite.

@@

\noindent
We first extend the \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
into a \SWT $T'$ over $\Sigma$ and $\hat\Omega$ (and the same semiring and label theory),
such that for all $s \in \Sigma^*$, and $u \in {\hat\Omega}^*$, 
$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simplify skips every symbol 
$b \in {\hat\Omega} \setminus \Omegai$ in input, 
with new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.

\noindent
Then, given an input word $s \in \Sigma^*$, 
we compute the \SWA $A_{T', s}$, using Proposition~\ref{prop:epsilon}.
This automaton is such that for all $t \in \T(\Omega)$, 
$A_{T', s}\bigl(\lin(t)\bigr) 
 = T'\bigl(s, \lin(t)\bigr) 
 = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
 = d(s, t)$.

>>>>

\item[--] a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
defining a measure $A: \T(\Omega) \to \Semiring$,
%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
%      $A : \Omega^* \to \Semiring$,
\item[--] an input word $s \in \Sigma^*$.
\end{description}
%
As explained in Section~\ref{sec:trees}, 
$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
or a labeled weighted CF grammar whose derivation trees are in $\T(\Omega)$, 
with weight defined by $A$.
The purpose of the transducer $T$ is to measure a distance between input and output words, 
\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Omegai^*$.
%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
%according to Definition~\ref{def:SWTA}, 
%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
%
%This actually is a slight generalization of CFG since 
%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
%hence parse trees %derivation trees
%are trees of $\T(\Omega)$.
%
%Another (more original) generalization is that the set of terminal symbols 
%$\Omega_0$ may be infinite.

@@

\noindent
We first extend the \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
into a \SWT $T'$ over $\Sigma$ and $\hat\Omega$ (and the same semiring and label theory),
such that for all $s \in \Sigma^*$, and $u \in {\hat\Omega}^*$, 
$T'(s, u) = T(s, u|_{\Omegai})$. The transducer $T'$ simply skips every symbol 
$b \in {\hat\Omega} \setminus \Omegai$, 
by the addition to the transition of $T$,
of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.

\noindent
Then, given an input word $s \in \Sigma^*$, 
we compute the \SWA $A_{T', s}$, using Proposition~\ref{prop:epsilon}.
This automaton is such that for all $t \in \T(\Omega)$, 
\[ 
   A_{T', s}\bigl(\lin(t)\bigr) 
 = T'\bigl(s, \lin(t)\bigr) 
 = T'\bigl(s, \lin(t)|_{\Omegai}\bigr) 
 = d(s, t).
\]
